```{r, label = setup, include = FALSE, cache = FALSE}
library(knitr)
opts_chunk$set(
              #  fig.path   = 'figure/'
                cache.path = 'cache/'
               , cache      = FALSE
               , dev        = c("svg", "tikz", "pdf", "png")[1]
               , fig.align  = 'center'
              # , fig.show   = 'asis'
              # , fig.width  = 6
              # , fig.height = 6
              # , out.width  = '0.9\\linewidth'
               , par        = TRUE
               , echo       = TRUE
               , results    = 'markup'
               , warning    = FALSE
               , message    = FALSE
               , comment    = NA
               )
options(
          replace.assign = TRUE
        , width          = 80
        , knitr.kable.NA = ''
        )

knit_hooks$set(par = function(before, options, envir){
if (before && options$fig.show!='none')
 par(mar=c(4, 4, 0.1, 0.1), cex.lab = 0.95, cex.axis =0.9, mgp = c(2, 0.7, 0), tcl = -0.3)
}, crop = hook_pdfcrop)

# Suppressing Error Messages in knitr
muffleError <- function(x,options) {}
knit_hooks$set(error=muffleError)
```

---
date: 2018-05-01
comments: false
slug: 2018-R-AARI
title: "An introduction to R for Agricultural Research"
categories:
- R
- Statistics
- Research
- Agriculture
---


## Introduction
R is a free, open-source programming language and software environment for statistical computing, bioinformatics, visualization and general computing. R provides a wide variety of statistical and graphical techniques, and is highly extensible. The latest version of R is 3.5.0 which can be obtained from [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/).

&nbsp;

## Regression Analysis
* Quantifying dependency of a normal response on quantitative explanatory variable(s)

```{r, label = df1}
Fert  <- seq(from = 0.3, to = 2.4, by = 0.3)
Yield <- c(10, 15, 30, 35, 25, 30, 50, 45)
df1   <- data.frame(Fert, Yield)
df1
```


```{r, label = p1}
library(ggplot2)
# if (!require("ggplot2")) install.packages("ggplot2")
p1 <- 
    ggplot(data = df1, mapping = aes(x = Fert, y = Yield)) + 
    geom_point() +
    labs(x = "Fertilizer", y = "Yield") +
    theme_bw()
print(p1)
```

```{r, label = fm1}
fm1 <- lm(formula = Yield ~ Fert, data = df1)
summary(fm1)
```


```{r, label = fm1anova}
fm1 <- lm(formula = Yield ~ Fert, data = df1)
anova(fm1)
```
