```{r, label = setup, include = FALSE, cache = FALSE}
library(knitr)
opts_chunk$set(
              #  fig.path   = 'figure/'
                cache.path = 'cache/'
               , cache      = FALSE
               , dev        = c("svg", "tikz", "pdf", "png")[1]
               , fig.align  = 'center'
              # , fig.show   = 'asis'
              # , fig.width  = 6
              # , fig.height = 6
              # , out.width  = '0.9\\linewidth'
               , par        = TRUE
               , echo       = TRUE
               , results    = 'markup'
               , warning    = FALSE
               , message    = FALSE
               , comment    = NA
               )
options(
          replace.assign = TRUE
        , width          = 80
        , knitr.kable.NA = ''
        )

knit_hooks$set(par = function(before, options, envir){
if (before && options$fig.show!='none')
 par(mar=c(4, 4, 0.1, 0.1), cex.lab = 0.95, cex.axis =0.9, mgp = c(2, 0.7, 0), tcl = -0.3)
}, crop = hook_pdfcrop)

# Suppressing Error Messages in knitr
muffleError <- function(x,options) {}
knit_hooks$set(error=muffleError)
```


```{r label = ARRI, echo = FALSE}
knitr::read_chunk("2018-05-09_ARRI.R")
```

---
date: 2018-05-09
comments: false
slug: 2018-R-AARI
title: "An introduction to R for Agricultural Research"
categories:
- R
- Statistics
- Research
- Agriculture
---


## Introduction
>**R** is a free, open-source programming language and software environment for statistical computing, bioinformatics, visualization and general computing. **R** provides a wide variety of statistical and graphical techniques, and is highly extensible. The latest version of **R** is 3.5.0 which can be obtained from [https://cran.r-project.org/bin/windows/base/](https://cran.r-project.org/bin/windows/base/).

&nbsp;


<h4><a href="https://cran.r-project.org/bin/windows/base/" class="badge badge-small"><i class="fa fa-file-pdf-o"></i>&nbsp;&nbsp;R</a> <a href="https://cran.r-project.org/bin/windows/base/">R-3.5.0 for Windows (32/64 bit)</a> &nbsp; </h4>

&nbsp;

<h4><a href="https://www.rstudio.com/products/rstudio/download/" class="badge badge-small"><i class="fa fa-file-pdf-o"></i>&nbsp;&nbsp;RStudio</a> <a href="https://www.rstudio.com/products/rstudio/download/">RStudio</a> &nbsp; </h4>


&nbsp;

<h4><a href="https://cran.r-project.org/doc/manuals/r-release/R-intro.pdf" class="badge badge-small"><i class="fa fa-file-pdf-o"></i>&nbsp;&nbsp;Manual</a> <a href="https://cran.r-project.org/doc/manuals/r-release/R-intro.pdf">An Introduction to R</a> &nbsp; </h4>

&nbsp;

<h4><a href="https://cran.r-project.org/doc/contrib/Short-refcard.pdf" class="badge badge-small"><i class="fa fa-file-pdf-o"></i>&nbsp;&nbsp;Ref Card</a> <a href="https://cran.r-project.org/doc/contrib/Short-refcard.pdf">R Reference Card</a> &nbsp; </h4>


&nbsp;

<h4><a href="https://github.com/myaseen208/R-Workshop/raw/master/2009NewYorkTimes.pdf" class="badge badge-small"><i class="fa fa-file-pdf-o"></i>&nbsp;&nbsp;New York Times</a> <a href="https://github.com/myaseen208/R-Workshop/raw/master/2009NewYorkTimes.pdf">New York Times</a> &nbsp; </h4>

&nbsp;

<h4><a href="https://github.com/myaseen208/R-Workshop/raw/master/2015NatureArticle.pdf" class="badge badge-small"><i class="fa fa-file-pdf-o"></i>&nbsp;&nbsp;Nature Article</a> <a href="https://github.com/myaseen208/R-Workshop/raw/master/2015NatureArticle.pdf">Nature Article</a> &nbsp; </h4>

&nbsp;


## Regression Analysis
> * Quantifying dependency of a normal response on quantitative explanatory variable(s)

&nbsp;

## Simple Linear Regression
> * Quantifying dependency of a normal response on a quantitative explanatory variable

&nbsp;

### Example

  Fertilizer (Kg/acre)  | Production (Bushels/acre)
  ------------| -------------
       100    |  70
       200    |  70
       400    |  80
       500    | 100
  

```{r, label = reg.df1}
```

```{r, label = reg.p1}
```

```{r, label = reg.fm1}
```


```{r, label = reg.fm1anova}
```
&nbsp;

### Example

  Fertilizer  | Yield
  ------------| -------------
       0.3    |  10
       0.6    |  15
       0.9    |  30
       1.2    |  35
       1.5    |  25
       1.8    |  30
       2.1    |  50
       2.4    |  45  
    

```{r, label = reg.df2}
```


```{r, label = reg.p2}
```

```{r, label = reg.fm2}
```


```{r, label = reg.fm2anova}
```

&nbsp;

## Multiple Linear Regression
> * Quantifying dependency of a normal response on two or more quantitative explanatory variables

&nbsp;

### Example

Fertilizer (Kg)  | Rainfall (mm) | Yield (Kg)
  ------------| -------------| -------------
       100    |  10   |  40
       200    |  20   |  50
       300    |  10   |  50
       400    |  30   |  70
       500    |  20   |  65
       600    |  20   |  65
       700    |  30   |  80

```{r, label = reg.df3}
```


```{r, label = reg.p3}
```

```{r, label = reg.fm3}
```


```{r, label = reg.fm3anova}
```


```{r, label = reg.fm3sbeta}
```


&nbsp;

## Polynomial Regression Analysis
> * Quantifying non-linear dependency of a normal response on quantitative explanatory variable(s)

### Example
> An experiment was conducted to evaluate the effects of different
levels of nitrogen. Three levels of nitrogen: 0, 10 and 20 grams per
plot were used in the experiment. Each treatment was replicated
twice and data is given below:
  
  Nitrogen    | Yield
  ------------| -------------
        0     |   5
        0     |   7
       10     |  15
       10     |  17
       20     |   9
       20     |  11


```{r, label = reg.df4}
```


```{r, label = reg.p4}
```

```{r, label = reg.fm4}
```


```{r, label = reg.fm4anova}
```

  