---
date: 2018-04-21
comments: false
slug: DiallelAnalysisR
title: "Diallel Analysis with R"
categories:
- Diallel Analysis
- R
---

The latest versin 0.1.1 of [DiallelAnalysisR](https://cran.r-project.org/web/packages/DiallelAnalysisR/index.html) is aviable on  [CRAN](https://cran.r-project.org/web/packages/DiallelAnalysisR/index.html). It performs Diallel Analysis with R using Griffing's and Hayman's approaches. Four different methods (1: Method-I (Parents + F1's + reciprocals); 2: Method-II (Parents and one set of F1's); 3: Method-III (One set of F1's and reciprocals); 4: Method-IV (One set of F1's only)) and two methods (1: Fixed Effects Model; 2: Random Effects Model) can be applied using Griffing's approach.

**DiallelAnalysisR** has functions to perform Diallel Analysis with R using Griffing's and Hayman's approaches. Four different Methods:

1. Method-I (Parents + F1's + reciprocals)
2. Method-II (Parents and one set of F1's)
3. Method-III (One set of F1's and reciprocals)
4. Method-IV (One set of F1's only)) 

and two Models:

1. Fixed Effects Model
2. Random Effects Model

can be applied using Griffing's approach.

## Installation
## Stable Version

Stable version of the package avialabe on [CRAN](https://cran.r-project.org/web/packages/DiallelAnalysisR/index.html) can be installed using:

```{r, eval = FALSE}
install.packages('DiallelAnalysisR', dependencies = TRUE)
```

## Development Version
Development version of the package avialabe on [Github](https://github.com/myaseen208/DiallelAnalysisR) can be installed using:

```{r, eval = FALSE}
if(!require("devtools")) install.packages("devtools")
devtools::install_github('myaseen208/DiallelAnalysisR', build_vignettes = TRUE)
```

## License
This package is free and open source software, licensed under GPL.
